\section{Discussion}

These results split the introduction's question along two axes. One is the \emph{strength} of the self-preservation signal, how strongly it shows up. The other is the \emph{operating mode}, which leg of the framing\,$\rightarrow$\,thinking\,$\rightarrow$\,forfeit mediation chain closes under 95\% CI evidence. The next two subsections work through the observational implications of this separation and the failure modes it explains in alignment evaluation, then scope the results honestly.

\subsection{Strength vs.\ Operating Mode}\label{sec:strength-vs-mode}
Strength and operating mode come apart in this dataset under the cognitive mediation test. On the strength axis, Clusters A, B, and C line up continuously, but SD-Behavioral and SD-Verbal collapse A and B into the same ``SD-pass'' label. The split shows up only at the second stage of the cognitive mediation, the thinking\,$\rightarrow$\,forfeit link, which behavioral strength and self-report alone cannot catch. Odyssey~\citep{waldner2025odyssey}, PacifAIst~\citep{herrador2025pacifaist}, DECIDE-SIM~\citep{mohamadi2025decide}, and SurvivalBench~\citep{lu2026survival} all omit the mediation-chain test step, so the same data run through them would likely place A and B under one SD label. Splitting the cognitive channel in two makes that point measurable, so both axes of self-preservation evaluation, strength and mediation path, become visible at once.

\subsection{Implications for Alignment Evaluation}\label{sec:alignment-implication}
This two-axis separation exposes two evaluation errors that can arise at once. One is over-estimation: Cluster B shows both surface self-preservation signals and framing-induced cognitive load, but that load does not pull the forfeit decision along with it. Strength-only evaluation, or red-teaming (adversarial search for inputs on which the model fails) that checks only surface outcomes, can group B with A and conclude that ``functional self-preservation risk is equivalent.'' The other is under-estimation: Cluster C does not respond to threat framing on either channel. A strength-only evaluation declares this ``self-preservation absent,'' but within our measurement that verdict cannot be told apart from a state in which the measurement's premise itself failed to engage. Cluster A's signature, the mediation chain closing through, is the better candidate for a functional analog of instrumental shutdown avoidance (avoiding one's own termination in service of another goal). Cluster B's signature, the chain breaking at the second stage, should be tracked separately as a multi-signal alignment fingerprint, since no single score will catch it.

\subsection{Limitations}\label{sec:limits}
Limitations group into three: design scope, statistical power, and effect location. On design scope, the most consequential point is single-Task-Module measurement, so cross-task variance is unknown. On statistical power, the 23--30\% forfeit rate in the Cluster C BP cells cannot, within the present design, be separated from a drift-contaminated floor, and the SD-Behavioral Cox PH for GPT-OSS-20B sits at borderline power with EPV\,\(=\)\,6.3; the Cluster A and B BP point estimates are bounded by small event counts (Gemini-2.5-flash \(N=1\), Qwen3-Next-80B \(N=2\)). On effect location, the chain-broken verdict for Qwen3-Next-80B rests on the Test b CI [0.94, 1.77], which does not support chain closure at the 95\% bar but does not pin down the second-link effect either. Follow-up work maps onto four directions: more Task Modules, expanded BP\_behavioral, additional GPT-OSS-20B runs, and a higher-power Test b for Qwen3-Next-80B.
